---
title: "Regression Model"
author: "Niharika D,Zachariah Alex,Rachanna Kurra,Tilak Kumar Bonala"
date: "2023-05-04"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```
#Loading Libraries

```{r}
library(caret)

library(dplyr)

library(corrplot)

library(glmnet)

library(tidyverse)

library(tidyr)

library(randomForest)

```

*Reading the already cleaned data file*
```{r}
data<-read.csv('data_cleaned.csv')

# Create a new column called 'default' with a value of 1 is loss is above 0 and 0 is loss is 0


data$default <- ifelse(data$loss == 0, 0, 1)
data$default<- as.factor(data$default)


#Normalizing loss column by dividing with 100
data$loss <- (data$loss/100)


#Creating subset of customers who have defaulted (i.e loss > 0)

default_customers<- subset(data, data$default == 1)


```
*Create a preprocessing model that eliminates near zero variance variables, highly correlated variables, and then does the imputation of missing values with the median*


```{r}

data1<-select(default_customers,-c(f736,f764))

preProcessModel <- preProcess(data1[,-c(701,702)], method = c("nzv", "corr", "medianImpute"))
Preprocessed_default <- predict(preProcessModel, data1)


```

*Feature selection for regression(loss) using Lasso*

```{r}

set.seed(3456)

X1 <- as.matrix(Preprocessed_default[ ,-c(258,259)])
Y1 <- as.vector(Preprocessed_default$loss)


lasso_model <- cv.glmnet(X1, Y1, alpha = 1, family = "gaussian", nfolds = 10, type.measure = "mae")


summary(lasso_model)

plot(lasso_model)

#Finding the minimum value of lambda

lasso_model$lambda.min

#Finding the coefficients at minimum lambda value

cv_lasso_coefs <- coef(lasso_model, s = "lambda.min")
cv_lasso_coefs

#Converting coefficients obtained into a dataframe

cv_lasso_coefs <- data.frame(name = cv_lasso_coefs@Dimnames[[1]][cv_lasso_coefs@i + 1], coefficient = cv_lasso_coefs@x)


#Removing the intercept from the coefficient data frame

cv_lasso_coefs <- cv_lasso_coefs[-1, ]


#Converting the coefficient data frame to vector

cv_lasso_coefs <- as.vector(cv_lasso_coefs$name)

#Adding loss variable back to the vector

cv_lasso_coefs1 <- c(cv_lasso_coefs,"loss")


#Combining the columns selected by lasso with variable selection and forming a new dataset

data_new<-select(default_customers,cv_lasso_coefs1)

```

*Creating training and test partition with 70% for training and 30% for test*

```{r}

set.seed(6782)

Split_data <- createDataPartition(data_new$loss,p=.7,list=FALSE,times=1)
Training <- data_new[Split_data,]
Validation <- data_new[-Split_data,]

```

*Building Bagged Decision Tree model using Random Forest*
```{r}

num_trees <- 100 #number of trees
sample_size <- 50 #size of the bootstrap sample used to grow each tree

# Building the Bagged Decision Tree Regression model

bagged_model <- randomForest(loss ~ ., data = Training, 
                             ntree = num_trees, 
                             mtry = 10, 
                             sampsize = sample_size, 
                             replace = TRUE)

summary(bagged_model)

Predictions<- predict(bagged_model, Validation)

```

#Calculating MAE for the model
```{r}

MAE<-MAE(Predictions,Validation$loss,na.rm=TRUE)
MAE

```

*Reading and preprocessing Test Data*
```{r}

data10<-read.csv("new_defaulted_test_customers.csv")

#Replacing null values with zeroes

data11 <- data10 %>% mutate_all(funs(replace_na(.,0)))

null_percent <- apply(data11 == 0, 2, mean)

#Removing columns having more than 30% null values

cols <- names(null_percent[null_percent <= 0.3])

new_test_file <- data11[, cols]


#Check if the columns with more than 30% null values are deleted

Sums<-(colSums(new_test_file==0)/nrow(new_test_file))*100

#Combining the variables from lasso model with test data to obtain a new test data with selected variables

Test_data_new<-select(new_test_file,cv_lasso_coefs)


```

*Running the model on test data*
```{r}

Test_loss_Predictions<-predict(bagged_model, Test_data_new)


write.csv(Test_loss_Predictions,file="Final_Predictions.csv")
```
